---
title: "p8105_hw1_jpv2129"
author: "Jake Vettoretti"
date: "2025-09-18"
output: github_document
---

## Problem 1

Loading the dataset
```{r}
library(tidyverse)
library(moderndive)
data("early_january_weather") #loads the dataset
```


Getting the number of rows/columns in the dataset and mean temp from the dataset.
```{r}
nrow(early_january_weather) #gets rows
ncol(early_january_weather) #gets columns
mean(with(early_january_weather, temp)) #calculating the mean temp
```

Dataset description:
There are 15 variables in the dataset.

The names of the variables are: origin, year, month, day, hour, temp, dewp, humid, wind_dir, wind_speed, wind_gust, 
precip, pressure, visib, and time_hour.

The important variables seem to be temp, dewp, humid, wind_dir, wind_speed, wind_gust, 
precip, pressure, and visib as these are the variables listed that mainly contribute to weather patterns.

There are 15 columns in the dataset and 358 rows (or observations).

The mean temp from the dataset is 39.58212 F.

Scatterplot:






