---
title: "p8105_hw1_jpv2129"
author: "Jake Vettoretti"
date: "2025-09-18"
output: github_document
---

## Problem 1

Loading the dataset
```{r}
library(tidyverse)
library(moderndive)
data("early_january_weather") #loads the dataset
```


Getting the number of rows/columns in the dataset and mean temp from the dataset.
```{r}
nrow(early_january_weather) #gets rows
ncol(early_january_weather) #gets columns
mean(with(early_january_weather, temp)) #calculating the mean temp
```

Dataset description:
There are 15 variables in the dataset.

The names of the variables are: origin, year, month, day, hour, temp, dewp, humid, wind_dir, wind_speed, wind_gust, 
precip, pressure, visib, and time_hour.

The important variables seem to be temp, dewp, humid, wind_dir, wind_speed, wind_gust, 
precip, pressure, and visib as these are the variables listed that mainly contribute to weather patterns.

There are 15 columns in the dataset and 358 rows (or observations).

The mean temp from the dataset is 39.58212 F.

Scatterplot:

```{r}
ggplot(early_january_weather, aes(x=time_hour,y=temp, color=humid)) + geom_point() #making the scatterplot with color points for humidity

ggsave("temp_vs_time_scatterplot.png") #saving scatterplot into project directory
```

Description of patterns in scatterplot:
The scatterplot shows temperature changes over time in January, with color indicating humidity. Overall, temperatures fluctuate daily but gradually increase from early to mid-January, starting mostly between 25–40°F through Jan 7th and more frequently rising to 45–55°F closer to January 14. 

Humidity increases with  warmer temperatures and spikes in mid-January when temperatures peak. Lower humidity values are more evenly spread and more common at cooler temperatures. 

## Problem 2

Creating the dataframe:

```{r}
problem2_df = #creating my own dataframe with conditions described.
  tibble(
    randsamp = rnorm(10), # random sample of size 10 from a standard Normal distribution
    logic = rnorm(10) > 0, # a logical vector indicating whether elements of the sample are greater than 0
    character = letters[1:10], # a character vector of length 10
    factor = factor(rep(c("A","B","C"), length.out = 10)) # a factor vector of length 10, with 3 different factor “levels”
  )

#taking means below
mean(problem2_df |>  pull(randsamp))    # works
mean(problem2_df |>  pull(logic))   # works
mean(problem2_df |>  pull(character))   # error 
mean(problem2_df |>  pull(factor))    # error 
```









